Learning and comparing functional connectomes between populations

### 2012-12-28 paper.tex
Iter on last part

\paragraph{Graph-topology metrics}
%
Functional connectivity graphs have been found to display specific
topological properties\footnote{In the neuroscience world, these
descriptions are group under the terms of ``graph theoretical
approaches'', however graph theory is an entire division of mathematics
and computer science that is concerned with much more than topology of
random graphs.} that are characteristic of small-word networks
\cite{stam2004,salvador2005,achard2006,bullmore2009}. These networks
display excellent transport properties: although they display relatively
small number of connections, any two regions of the brain are well
connected. Another interesting consequence of their specific topology is
the resilience it gives to the system to attacks such as resulting from
brain lesions \cite{achard2006}. This overall structure of
functional-connectivity graphs can be summarized by a few metrics, such
as the average path length between any two nodes or the local clustering
coefficients \cite{rubinov2010}. Given that pathologies without a well-defined focus, such
as schizophrenia, are thought to have a global impact on brain
connectivity \cite{liu2008,bassett2008}, it is interesting to use the
graph-topology metrics to perform inter-subject comparison. Such an
approach is appealing as it is not subject to multiple comparison issues.
However, it has been criticized as giving a fairly unspecific
characterization of the brain and being fragile to noise
\cite{ioannides2007}. Another caveat is that correlation matrices display
small-word properties such as local clustering by construction, as if two
nodes are strongly correlated to a third, they are highly likely to be
correlated to each other \cite{zalesky2012}. This observation highlight
the need for well defined null-hypothesis \cite{zalesky2012,rubinov2011},
but also for controlled recovery of brain functional connectivity going
beyond empirical correlation.

@@

\section{Functional and effective connectivity}

Effective connectivity does not have to be causal => but a precise mathematical
definition of the relationship => BCI workshop reports

Fucntional connectivity is a statistical relationship, instead of precise mathematical
relationship

@@

Cross validation???

>>>>

\paragraph{Graph-topology metrics}
%
Functional connectivity graphs have been found to display specific
topological properties\footnote{In the neuroscience world, these
descriptions are group under the terms of ``graph theoretical
approaches'', however graph theory is an entire division of mathematics
and computer science that is concerned with much more than topology of
random graphs.} that are characteristic of small-word networks
\cite{stam2004,salvador2005,achard2006,bullmore2009}. These networks
display excellent transport properties: although they display relatively
small number of connections, any two regions of the brain are well
connected. Another interesting consequence of their specific topology is
the resilience it gives to the system to attacks such as resulting from
brain lesions \cite{achard2006}. This overall structure of
functional-connectivity graphs can be summarized by a few metrics, such
as the average path length between any two nodes or the local clustering
coefficients \cite{rubinov2010}. Given that pathologies without a well-defined focus, such
as schizophrenia, are thought to have a global impact on brain
connectivity \cite{liu2008,bassett2008}, it is interesting to use the
graph-topology metrics to perform inter-subject comparison. Such an
approach is appealing as it is not subject to multiple comparison issues.
However, it has been criticized as giving a fairly unspecific
characterization of the brain and being fragile to noise
\cite{ioannides2007}. Another caveat is that correlation matrices display
small-word properties such as local clustering by construction, as if two
nodes are strongly correlated to a third, they are highly likely to be
correlated to each other \cite{zalesky2012}. This observation highlight
the need for well defined null-hypothesis \cite{zalesky2012,rubinov2011},
but also for controlled recovery of brain functional connectivity going
beyond empirical correlation matrices.

@@

\section{Beyond correlation, effective connectivity?}

All the approaches that we have presented in this review are based on
second-order statistics of the signal, in other words correlation
analysis. Traditionally, these are defined as \emph{functional
connectivity}, defined as ``temporal correlations between remote
neurophysiological events'' \cite{friston1994}, and opposed to
\emph{effective connectivity}, \emph{i.e.}\ ``the influence one neural
system exerts over another'' \cite{friston1994}. To conclude this review,
we would like to bridge the gap between these concepts, which in our eyes
should be seen as a continuum rather than an opposition (this opinion is
also expressed in \cite{mcintosh2010}).

A first step to move from purely descriptive statistics to modeling with
functional connectivity analysis is to consider a correlation matrix as a
Gaussian graphical model, \emph{i.e.} a well-defined probabilistic model
that describes observed correlations in terms of an independence
structure and conditional relations \cite{lauritzen1996,varoquaux2011}.
In such settings, the inverse covariance graph or the partial
correlations are a measure of influence from one node to another, albeit
undirected.

@@

Cross validation
